---
title: "Transform"
author: "Peter Desmet"
format: html
editor: visual
---

## Read data

```{r}
#| echo: false
library(readr)
library(here)
library(dplyr)
library(frictionless)
```

```{r}
dwc_event <- read_tsv(here::here("data", "dwc", "event.txt"))
dwc_occurrence <- read_tsv(here::here("data", "dwc", "occurrence.txt"))
dwc_mof <- read_tsv(here::here("data", "dwc", "measurementorfacts.txt"))
```

## Approach 1

### Create data package

```{r}
dp1 <- create_package()
```

### Process events (including occurrences)

Process event data:

```{r}
dp1_event1 <-
  dwc_event %>%
  select(-id) %>%
  mutate(parentEventID = NA) %>% # Original parentEventID not relevant
  mutate(eventType = "visit") %>%
  relocate(eventType, eventID, parentEventID, .before = NULL)
```

Process occurrence data:

```{r}
dp1_event2 <-
  dwc_occurrence %>%
  select(-id) %>%
  rename(parentEventID = eventID) %>% # The (higher) eventID is a parentEventID
  rename(eventID = occurrenceID) %>% # The occurrenceID is an eventID
  mutate(eventType = "occurrence") %>%
  relocate(eventType, eventID, parentEventID, .before = NULL)
```

Combine both into single table:

```{r}
dp1_event <- bind_rows(dp1_event1, dp1_event2)
```

Create Table Schema, incl. relationships:

```{r}
dp1_event_schema <- create_schema(dp1_event) # Creates list of fields
keys <- list(
  primaryKeys = list(c("eventID")),
  uniqueKeys = list(c("eventID")),
  foreignKeys = list(c(
    list(
      fields = list(c("parentEventID")),
      reference = list(
        resource = "event",
        fields = list(c("eventID"))
      )
    )
  ))
)
dp1_event_schema <- append(dp1_event_schema, keys)
```

Add data as resource:

```{r}
dp1 <-
  dp1 %>%
  add_resource(
    resource_name = "event",
    data = dp1_event,
    schema = dp1_event_schema
  )
```

### Process Measurements or facts

Process data:

```{r}
dp1_mof <- dwc_mof # No transformation required
```

Create schema + relationships:

```{r}
dp1_mof_schema <- create_schema(dp1_mof)
keys <- list(
  primaryKeys = NULL,
  uniqueKeys = NULL, # No measurementID defined
  foreignKeys = list(c(
    list(
      fields = list(c("id")),
      reference = list(
        resource = "event",
        fields = list(c("eventID"))
      )
    )
  ))
)
dp1_mof_schema <- append(dp1_mof_schema, keys)
```

Add data as resource:

```{r}
dp1 <-
  dp1 %>%
  add_resource(
    resource_name = "measurementorfact",
    data = dp1_mof,
    schema = dp1_mof_schema
  )
```

### Write data package

```{r}
write_package(dp1, here::here("data", "dwc-dp1"))
```

## Approach 2

### Create data package

```{r}
dp2 <- create_package()
```

### Process events

Process data:

```{r}
dp2_event <-
  dwc_event %>%
  select(-id) %>%
  mutate(parentEventID = NA) %>% # Original parentEventID not relevant
  mutate(eventType = "visit") %>%
  relocate(eventType, eventID, parentEventID, .before = NULL)
```

Create Table Schema, incl. relationships:

```{r}
dp2_event_schema <- create_schema(dp2_event)
keys <- list(
  primaryKeys = list(c("eventID")),
  uniqueKeys = list(c("eventID")),
  foreignKeys = list(c(
    list(
      fields = list(c("parentEventID")), # These are not used in this dataset
      reference = list(
        resource = "event",
        fields = list(c("eventID"))
      )
    )
  ))
)
dp2_event_schema <- append(dp2_event_schema, keys)
```

Add data as resource:

```{r}
dp2 <-
  dp2 %>%
  add_resource(
    resource_name = "event",
    data = dp2_event,
    schema = dp2_event_schema
  )
```

### Process occurrences

Process data:

```{r}
dp2_occurrence <-
  dwc_occurrence %>%
  select(-eventID) %>% # Duplicate of id
  relocate(id, occurrenceID, .before = NULL)
```

Create Table Schema, incl. relationships:

```{r}
dp2_occurrence_schema <- create_schema(dp2_occurrence)
keys <- list(
  primaryKeys = list(c("occurrenceID")),
  uniqueKeys = list(c("occurrenceID")),
  foreignKeys = list(c(
    list(
      fields = list(c("id")),
      reference = list(
        resource = "event",
        fields = list(c("eventID"))
      )
    )
  ))
)
dp2_occurrence_schema <- append(dp2_occurrence_schema, keys)
```

Add data as resource:

```{r}
dp2 <-
  dp2 %>%
  add_resource(
    resource_name = "occurrence",
    data = dp2_occurrence,
    schema = dp2_occurrence_schema
  )
```

### Process Measurements or facts

Process data:

```{r}
dp2_mof <- dwc_mof # No transformation required
```

Create schema + relationships:

```{r}
dp2_mof_schema <- create_schema(dp2_mof)
keys <- list(
  primaryKeys = NULL,
  uniqueKeys = NULL, # No measurementID defined
  foreignKeys = list(c(
    list(
      fields = list(c("id")),
      reference = list(
        resource = "event",
        fields = list(c("eventID"))
      )
    )
  ))
)
dp2_mof_schema <- append(dp2_mof_schema, keys)
```

Add data as resource:

```{r}
dp2 <-
  dp2 %>%
  add_resource(
    resource_name = "measurementorfact",
    data = dp2_mof,
    schema = dp2_mof_schema
  )
```

### Write data package

```{r}
write_package(dp2, here::here("data", "dwc-dp2"))
```

## 
